---
title: "Assignment 1 - Language Development in ASD - part 3"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/ra99bdvm6fzay3g/demo_test.csv?dl=0
* Utterance Length data: https://www.dropbox.com/s/uxtqqzl18nwxowq/LU_test.csv?dl=0
* Word data: https://www.dropbox.com/s/1ces4hv8kh0stov/token_test.csv?dl=0


```{r}
#------ Load library ------ 
library(data.table)
library(ggplot2)
library(stringi)
library(stringr)
library(tidyverse)
library(zoo)
library(plyr)
library(dplyr)
library(lmerTest)
library(lme4)
library(MuMIn)
library(modelr)
library(caret)
library(ModelMetrics)

#Cleaning the data

#------ Load data ------
wd=getwd()
setwd(wd)

demo_test = read.csv("demo_test.csv")
LU_test = read.csv("LU_test.csv")
token_test = read.csv("token_test.csv")

#------ fix variables for merge ------
#Standardizing variable names
demo_test= setnames(demo_test,"Visit", "VISIT")
demo_test= setnames(demo_test, "Child.ID", "SUBJ")

#Retrieve digits to make visit a number
demo_test$VISIT= str_extract(demo_test$VISIT, "\\d")
LU_test$VISIT= str_extract(LU_test$VISIT, "\\d")
token_test$VISIT= str_extract(token_test$VISIT, "\\d")

#Removes punctuation from SUBJ
LU_test$SUBJ  = gsub("\\.", "", LU_test$SUBJ)
demo_test$SUBJ  = gsub("\\.", "", demo_test$SUBJ)
token_test$SUBJ  = gsub("\\.", "", token_test$SUBJ)

#Translating variable names to Human
demo_test= setnames(demo_test, "MullenRaw", "nonverbalIQ")
demo_test= setnames(demo_test, "ExpressiveLangRaw", "verbalIQ")

#------ Merge data ------
#Make subsets of relevant variables
demo_test_sub = select(demo_test, c("SUBJ", "VISIT", "Ethnicity", "Diagnosis", "Gender", "Age", "ADOS", "nonverbalIQ", "verbalIQ")) 
LU_test_sub = select(LU_test, c("SUBJ", "VISIT", "MOT_MLU", "MOT_LUstd", "CHI_MLU", "CHI_LUstd"))
token_test_sub = select(token_test, c("SUBJ", "VISIT", "types_MOT", "types_CHI", "tokens_MOT", "tokens_CHI"))

#Merge above subset
datamerge1 = merge(demo_test_sub,LU_test_sub,by= c("SUBJ", "VISIT"))
datamerge2 = merge(datamerge1, token_test_sub, by=c("SUBJ", "VISIT"))

#------ Edit merged data ------ 
#For these five variables select rows where VISIT==1
small_sub=subset(select_(datamerge2,"SUBJ", "VISIT", "ADOS", "nonverbalIQ", "verbalIQ"),VISIT==1)

#merge the dataframes. This gives double columns for the variables in question. Except for visit, columns with _old should be deleted and the new kept
datamerge3=merge(small_sub,datamerge2,by= c("SUBJ"), suffixes = c("","_old"))

#Remove unnecessary columns
datamerge3=datamerge3[,!names(datamerge3) %in% c("VISIT","ADOS_old","nonverbalIQ_old","verbalIQ_old")]

#rename VISITS_old to VISIT
datamerge3=dplyr::rename(datamerge3, VISIT =VISIT_old)


#anomizing subjects by making them numbers
datamerge3$SUBJ = as.numeric(as.factor(datamerge3$SUBJ))

#Make VISIT numeric
datamerge3$VISIT = as.numeric(datamerge3$VISIT)

#Translate levels to Human
datamerge3$Gender = mapvalues(datamerge3$Gender, "1","M")
datamerge3$Diagnosis = mapvalues(datamerge3$Diagnosis, c("A", "B"),c("ASD", "TD"))

#------ Write CSV file with the test data ------
write.csv(datamerge3, file = "final_test_data.csv", row.names = FALSE)


```


### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data () and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the code again and apply it to Assignment2TrainData1.csv)
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the functions rmse() and predict() )
- create the test dataset (apply the code from assignment 1 part 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: time to reuse "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())

```{r}
#------ Load the data ------
test_data = read.csv("final_test_data.csv")
data = read.csv("finaldata.csv")

#------ Use model on both test and training data ------
# removed gender because test data only had males
#Model on training data
m6=lmer(CHI_MLU~Diagnosis*VISIT+ADOS+MOT_MLU+(1+SUBJ|VISIT),data = data)
summary(m6)
modelr::rmse(m6, data) #note that both modelr and modelmetrics have an rmse function

#Model on test data
Predicted=predict(m6, type='response')
ModelMetrics::rmse(test_data$CHI_MLU, Predicted)




#m7=lmer(CHI_MLU~Diagnosis*VISIT+ADOS+MOT_MLU+(1+SUBJ|VISIT),data = test_data)
#modelr::rmse(m7, data_test)

```


Reporting


### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).
- Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold)
- Report the results and comment on them.

- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: How would you go comparing the performance of the basic model and the cross-validated model on the testing set?
- Bonus Question 2: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 3: compare the cross-valida
ted predictive error against the actual predictive error on the test data

```{r}
# ------ Merging train data and test data ------
#test_data$SUBJ = test_data$SUBJ + 61
full_data = rbind(data, test_data)


#------ Fold data ------
#Create fold with unique SUBJ
folds=createFolds(unique(full_data$SUBJ), k = 4)

#--------Create matarix to store results--------
cv_empty=matrix(ncol=3, nrow=4)


#----------------Loop----------------
#Indexing, reset when using the loop again
n=1

for (k in folds){
  #------ Split into training and test data ------ 
  #Create training dataset, data not in fold k
  data_train=subset(full_data,!(SUBJ %in% k))
  #Create test dataset, data in fold k
  data_test=subset(full_data,SUBJ %in% k)
  
  #------ train model - apply model to data_train ------
  model= lmer(CHI_MLU~Diagnosis*VISIT+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),data_train)
  #Make predictions based on model
  Prediction=predict(model, type='response')
  #Prints the prediction against the test data
  results_train=ModelMetrics::rmse(data_train$CHI_MLU, Prediction)
  
  #------ test the model - test model on data_test (last quarter) ------
  #Make predictions based on modeVIS
  Prediction_test=predict(model, data_test, type='response', allow.new.levels = TRUE)
  #
  results_test=ModelMetrics::rmse(data_test$CHI_MLU, Prediction_test)
  
  #------ save the performance ------ 
  #Fills out the kth row in cv_empty with iteration and rmse's
  cv_empty[n,]=cbind(n,results_train,results_test)
  n=n+1
  }

#------ Make the matrix into a data frame ------ 
cv_results=data.frame(cv_empty)
#Rename variables 
cv_results=setnames(cv_results,"X1", "iteration")
cv_results=setnames(cv_results,"X2", "train")
cv_results=setnames(cv_results,"X3", "test")


#------ Model on full data ------
m7= lmer(CHI_MLU~Diagnosis*VISIT+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),full_data)
summary(m7)
modelr::rmse(m7, full_data)


# Funilly enough, diagnosis and ADOS are correlated hence we use an interaction effect in the model
ggplot(data, aes(Diagnosis,ADOS)) + geom_boxplot()

```


### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.
(Tip: recreate the equation of the model: Y=Intercept+BetaX1+BetaX2, etc; input the average of the TD group  for each parameter in the model as X1, X2, etc.).

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r}
#------ Data ------
#Creating a dataset without bernie (Bernie is number 63)
no_bernie_data = subset(full_data, (SUBJ != 63))
#ONLY bernie
lonely_bernie_data = subset(full_data, (SUBJ == 63))

#----- Create model and make average TD child at each visit ------
# creating final model
model_full = lmer(CHI_MLU~Diagnosis*VISIT+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),no_bernie_data)

#Get betas
summary(model_full)

# getting means of variables for the TD kids, used for input as "x" in model
mean(full_data$ADOS[full_data$Diagnosis == "TD"])
mean(full_data$MOT_MLU[full_data$VISIT==6])
mean(full_data$verbalIQ[full_data$Diagnosis == "TD"])

# Calculating predicted mean of TD kids per visit
MLU_visit1 = 0.71 + 0.72*1 + 0.06*1 + 0.08*1 - 0.02*0.78 + 0.34*3.56 + 0.06*19.81 + 0.24*1*1

MLU_visit2 = 0.71 + 0.72*1 + 0.06*2 + 0.08*1 - 0.02*0.78 + 0.34*3.76 + 0.06*19.81 + 0.24*1*2

MLU_visit3 = 0.71 + 0.72*1 + 0.06*3 + 0.08*1 - 0.02*0.78 + 0.34*3.92 + 0.06*19.81 + 0.24*1*3

MLU_visit4 = 0.71 + 0.72*1 + 0.06*4 + 0.08*1 - 0.02*0.78 + 0.34*4.04 + 0.06*19.81 + 0.24*1*4

MLU_visit5 = 0.71 + 0.72*1 + 0.06*5 + 0.08*1 - 0.02*0.78 + 0.34*4.13 + 0.06*19.81 + 0.24*1*5

MLU_visit6 = 0.71 + 0.72*1 + 0.06*6 + 0.08*1 - 0.02*0.78 + 0.34*4.09 + 0.06*19.81 + 0.24*1*6

# making a list of MLU of TD kids at the different visits as calculated above.
MLU_list = c(MLU_visit1, MLU_visit2, MLU_visit3, MLU_visit4, MLU_visit5, MLU_visit6)


#------ Bernie vs TD & Bernie vs Prediction ------

#getting the difference in MLU between Bernie and the TD per visit
lonely_bernie_data[,12] - MLU_list #Worse than average TD

#getting the difference between Bernie and his predicted score
Prediction=predict(model_full, lonely_bernie_data[6,], allow.new.levels = TRUE)
lonely_bernie_data[6,12]-Prediction #Better than expected






# bernie MLU at visit 6
#bernie_MLU_visit6 = 0.71 + 0.72*0 + 0.06*6 + 0.08*1 - 0.02*7 + 0.34*3.82 + 0.06*28 + 0.24*0*6

```


[HERE GOES YOUR ANSWER]

### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!


