---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Riccardo Fusaroli"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}
#------ WD ------
wd=getwd()
setwd(wd)

#------ Load library ------ 
library(simr)
library(MASS)

#------ favorite model ------
model_full = lmer(CHI_MLU~Diagnosis*VISIT+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),full_data)

model_sub = lmer(CHI_MLU~Diagnosis+VISIT+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),full_data)

#------ Power simulations ------
power_VISIT= powerSim(model_sub, fixed("VISIT"), nsim= 200)
power_VISIT
#Power = 100, confidence interval 98.17 - 100, unknown test with effect size = 0.19
 
power_Diagnosis= powerSim(model_sub, fixed("Diagnosis"), nsim= 200)
power_Diagnosis
#Power = 14.50, confidence interval 9.93 - 20.16, Likelihood ratio

power_Gender= powerSim(model_sub, fixed("Gender"), nsim= 200)
power_Gender
#Power = 14.00, confidence interval 9.51 - 19.59, Likelihood ratio

power_ADOS= powerSim(model_sub, fixed("ADOS"), nsim= 200)
power_ADOS
#Power = 11.00, confidence interval 7.02 - 16.18, unknown test with effect size = -0.015

power_MOT_MLU= powerSim(model_sub, fixed("MOT_MLU"), nsim= 200)
power_MOT_MLU
#Power = 100, confidence interval 98.17 - 100, unknown test with effect size = 0.34

power_verbalIQ= powerSim(model_sub, fixed("verbalIQ"), nsim= 200)
power_verbalIQ
#Power = 100, confidence interval 98.17 - 100, unknown test with effect size = 0.063

power_INT = powerSim(model_full, fixed("VISIT*Diagnosis"), nsim= 200)
power_INT
#Power = 0, confidence interval 0 - 1.83, unknown test

# Hvad betyder power outputtet for de enkelte fixed effects? Hvad kan vi bruge det til?
# Power fra interaction er 0%
```


### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

# minimum effects sizes for each fixed effect

fixef(model_full)["VISIT"] = 0.1
fixef(model_full)["DiagnosisTD"] = 0.1
fixef(model_full)["GenderM"] = 0.1
fixef(model_full)["ADOS"] = 0.1
fixef(model_full)["MOT_MLU"] = 0.1
fixef(model_full)["verbalIQ"] = 0.1

# assesing power for each fixed effect by SUBJ
powerCurveV = powerCurve(model_sub, fixed("VISIT"),along="SUBJ", nsim=5)
powerCurveV

powerCurveD = powerCurve(model_sub, fixed("DiagnosisTD"),along="SUBJ", nsim=5)
powerCurveD

powerCurveVD = powerCurve(model_full, fixed("VISIT:DiagnosisTD"),along="SUBJ", nsim=5)
powerCurveVD

powerCurveG = powerCurve(model_full, fixed("GenderM"),along="SUBJ", nsim=5)
powerCurveG

powerCurveA = powerCurve(model_full, fixed("ADOS"),along="SUBJ", nsim=5)
powerCurveA

powerCurveM = powerCurve(model_full, fixed("MOT_MLU"),along="SUBJ", nsim=5)
powerCurveM

powerCurveVer = powerCurve(model_full, fixed("verbalIQ"),along="SUBJ", nsim=5)
powerCurveVer

# simulating new data
participants = 200
visits = 6
model = model_full


#------ analysis using simulated data ------
sim_data = createNewData(participants, visits, model)


# But Ricardos function only produces a new dataset with Visit, SUBJ, Diagnosis, CHI_MLU

# WE GOT THIS FAR
model_sim = model_full = lmer(CHI_MLU~Diagnosis*Visit+Gender+ADOS+MOT_MLU+verbalIQ+(1+VISIT|SUBJ),sim_data)



### Riccardo's clumsy function to simulate new participants
### TO DO points are only notes for myself, so not part of the assignment

createNewData <- function (participants,visits,model){
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(Visit=1:visits,SUBJ=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$SUBJ]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma = Sigma.u)
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$Visit + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}



```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.

```{r}

# Outputs of power analysis
powerCurveV # 100% power at 30 particpants
powerCurveD # 0% power at 30 particpants
powerCurveG # 0% power at 30 particpants
powerCurveA # 0% power at 30 particpants
powerCurveM # 60% power at 30 particpants
powerCurveVer # 100 power at 30 particpants

# All in all, it only makes sense to run the study on 30 participants if we were interested in only visit, gender and verbal IQ. Since we're interested in the effect of diagnosis, the study should't be run.

```




